Submodular Optimization for Control of Prosumer Networks

### 2016-04-19 Article.tex
update

In this paper, we study the case of a network composed of entities called prosumers. These agents have the ability to both consume and produce according to external conditions. We first design a model and simplify the underlying dynamic of such network using a second order coupled oscillators network model. Under some conditions, the system synchronizes to a common frequency. However, in case of perturbations in the power distribution, the system might lose synchrony and require control to bring it back to the stable state. In this situation, control can be seen as energy  absorbed or injected in the system at specific locations. Moreover, the power outputs of the prosumers are susceptible to change such that loads and generators are not fixed. In this context, it is important to select the right subset of nodes that yields, on average, the cheapest control in terms of energy. We propose and validate an algorithm based on a submodular optimization for chosing the driver nodes.
\end{abstract}

@@

The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems. Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility. Within these systems, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication. Agents responsible for generation and load portfolio managment, the so-called prosumers \cite{Parag2016} because they both produce and consume electricity, will provide services (generation, load shedding, frequency regulation) against remuneration \cite{Gensollen2014}. Insure grid stability within this uncertain context appears as a complex task.

Several studies highlighted the need for increasing the storage capacity of the system. Some of them even consider electric vehicles as moving capacities that could be used for frequency regulation. In the case of fixed storage devices, the locations where they should be installed is an important question since it impacts the performances of the system. In a very static and top-down architecture, where loads and generators are fixed, this question is far from trivial. But in a smart grid scenario with bi-directional flows and agents that produce and consume depending on weather conditions, this becomes even more challenging. In this direction, \cite{Gkatzikis2015} studies the placement of storage equipments within microgrids in a collaborative scheme. The authors solve an interesting optimization problem mixing placement-dimensioning of storage with utilization. Moreover, \cite{Gkatzikis2015} uses a Nash bargaining framework to determine how prosumers should share the costs and benefits of the storages. However, \cite{Gkatzikis2015} does not take into account the grid dynamic as well as its controllability. Control of smart grid systems has been recently studied in \cite{Farraj2015}, where the authors use linear-quadratic optimal control theory in order to compute DER outputs. Control inputs rely on information collected by sensors and phasor measurement units and a communication network is used for sharing the information. The performances are then studied under practical limitations such as latency, sampling rate, or signal-to-noise ratio. The objective of \cite{Farraj2015} is the control of the system when perturbations occur. The control signals are applied to the DER, meaning that \cite{Farraj2015} does not consider a system where storage would be present and used for control.

@@

In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes. In such a scenario with a high penetration of renewables, perturbations in the power distributions of the prosumers are very likely. Re-balancing the production and the consumption is then necessary and requires that we have the capability to control the grid's dynamic. The present work differs from \cite{Gkatzikis2015} and \cite{Farraj2015} in that we study the placement of the storages from a controllability perspective in a stochastic context due to DER. Moreover, we aim at models and methods that could be scallable to some extent.

We are given a network of prosumers, and we have to find both the number and locations of the storages that should be deployed. Once this choice is made, we have to stick with it afterwards whatever the perturbations that may occur. We are looking for the smallest set of storages, as well as their locations, that will require small control energy. Since we do not have any prior on the future perturbations, we minimize the average control energy required to move the system around the state space.

@@

The power grid dynamic is represented by a coupled oscillator network. Each node, representing a prosumer, is modeled as an oscillator with a phase angle and a natural frequency that depends on its production/consumption. The oscillators are interconnected such that their frequencies are impacted by the ones of their neighbors. As we will see, the phase angles dynamics can be written as a system of coupled second order differential equations \cite{Filatrella2008}.

@@

In this section, we introduce some key notions related to the controllability of complex networks. We consider a graph $G=(V,E)$ of N nodes, where V is the vertex set and E is the edge set. The topology of this network can be encoded through the adjacency matrix M, which elements $m_{ij} \in \{0,1\}$ indicate whether i and j are connected. If the edges are weighted we can set $m_{ij} = w_{ij}$ where $w_{ij}$ is the weight of edge$(i,j)$, and zero if there is no edge. Graphs are often used to model complex interactions inside a population. Let us attribute to each node i the variable $ x_i $ that represents the state of node i. Depending on the model, this state can represent an opinion, a physical quantity, a probability and so on. If there is an edge between i and j, we might expect some interaction that would eventually change the values of $x_i$ and $x_j$. Such a dynamic is usually described with a system of differential equations. If we assume a linear first order dynamic, this can be written as $ \dot{X}(t) = A X(t) $, where $X(t)=\{x_0(t),...,x_{N-1}(t)\}$ is the vector of node states at time t, and matrix A encodes both the dynamics and the topology of the system (A can be, among other, the weighted adjacency matrix of the underlying graph).

If there is no action from the outside, the system state will evolve according to $ X(t) = X_0 e^{At} $ where $X_0$ is the initial state of the system. Imagine now, that this dynamic could be somehow influenced by injecting some signals. More precisely, lets assume that we can inject these signals in a subset of the nodes such that the dynamics becomes $ \dot{ X }(t) = A X(t) + B u(t) $, where the matrix B indicates which nodes receive the signals and $ u(t) $ is the vector of inputs at time t. Assume that the system is in some initial state $X_0$ and we aim at bringing it to some final state $X_T$ in some amount of time T. Control can be seen as finding the sequence of inputs $\{u(t_0),...u(t_T)\} $ that would do it given the dynamics A. Furthermore, we say that the system is controllable in T steps if it can be steered from any initial state $ X_0 $ to any final state $ X_T $ through a sequence of control inputs.

It was shown by Kalman in 1963 that a linear system $(A,B)$ is controllable if the controllability matrix $ C = [B,AB,A^2B,...,A^{N-1}B] $ has full rank, that is, $rank(C) = N$. Nevertheless, when studying the controllability of non trivial networks, using the Kalman's criterion is not a simple task since matrix C becomes huge. Furthermore, because $A^k\longrightarrow 0$ when $k\longrightarrow \infty $, numeric precision becomes quickly an issue when computing the rank of C. A different approach to study controllability in complex network was introduced later on by Lin \cite{1100557}. This relates the search of the minimum driver set to a maximum matching in a bipartite graph, which can be done with the Hopcroft-Karp algorithm in $ O(|E|\sqrt{|V|}) $ in the worst case. This requires that the system $(A,B)$ is a structured system, meaning that all elements in A and B are either fixed zeros or independant free parameters. In such a situation, structural control is extremely powerfull since it allows to discuss network control easily even if the exact weights of the edges are unknown. Note however that the independance of the parameters is a key assumption. If A is the adjacency matrix of an undirected network for instance, independance is not verified (since A is symetric). In this case, one should use different tools to study controllability \cite{Yuan2014}.

The fact that a system is controllable or not is not the only criterion of interest. Indeed, if it is controllable, the amount of control energy required also appears as a key criterion. Actually, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $X_0$ to $X_T$. Among all these possibilities, optimal control is devoted to find the sequence that minimizes some cost function that can depend on the sates of the system $X(t)$, the control inputs $u(t)$, and the final state $X_T$. In the following, these control inputs will be used to model the actions of storage devices in a power grid. We are therefore particularly interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}. Let $ \mathcal{E} = \int_{0}^{T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system. It can be shown \cite{Liu2015} that the control inputs that minimize $ \mathcal{E} $ can be written as :

@@

where $ \nu_f = X_T - X_0 e^{AT} $ is the difference between the desired final state $X_T$ and the final free state $X_0 e^{AT} $, and $W(t) = \int_{0}^{t} e^{A \tau}BB^{T} e^{A^{T} \tau} d\tau $ is called the Gramian matrix of the system. It can also be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace. Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :

@@

where $ K_{Di} $ is the dissipation constant of entity i.

Furthermore, it also accumulates kinetic energy at a rate :

@@

where $ I_{i}$ is the moment of inertia of entity i. For simplicity, we consider that all entities have the same dissipation constants($K_D$) and moment of inertia (I).

@@

with $ P_{ij}^{MAX} $ being the maximum capacity of the line $(i,j)$. Each entity i is then described by a power balance equation of the type :

@@

where $ P_{S,i} $ is the power of an ideal source or sink at node i. By substituting equations \ref{eq:diss}, \ref{eq:acc}, and \ref{eq:trans} into equation \ref{eq:bilan} and re-arranging the terms, we obtain the following non-linear coupled system of equations :

@@

where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $. In order not to overload the equations, we simplify the constant term $ \frac{K_D \Omega}{I} $ by working in a rotating frame such that $ \psi_i=\frac{P_{S,i}}{I \Omega} $.

@@

With transition matrix A :

@@

On the other hand, the power that flows on line $ (i,j) \in E$ can be written as $ P_{i \longrightarrow j} = -P_{ij}^{MAX} sin\left( \theta_j(t) - \theta_i(t) \right) $. If $ \theta_j(t) - \theta_i(t) \sim 0 $, the flow constraints at time t are straightforward :

@@

Constraints \ref{synchro_constraint} and \ref{zero_sum_constraint} ensure that the system is able to synchronize at $\Omega$ without overloading lines.

@@

Where $ \Lambda(0) $ is the vector of initial levels in the batteries. Obviously, $\Lambda(t)$ has to stay within possible bounds : $ \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{MAX} $. Which can be written in terms of the control inputs :

@@

Recall that $\mathcal{E}_{min}$ (see eq. \ref{eq:energy}) depends on the initial and final states as well as on the inverse of the gramian matrix W, which only depends on A and B. W can thus be used to obtain information about the average control energy required to move the system in the state space. In the prosumer network that we consider, generators and loads ($\Psi$) are suceptible to change, meaning that initial ($X_0$) and final ($X_f$) states might also vary. In such a scenario, we prefer to aim at good performances on average, rather than very good performances in few specific cases and bad performances in all other situations.

@@

\begin{figure}
	\label{fig:submodular}
	\begin{center}
	\includegraphics[scale=.25]{./figure_1/figure_1}
	\caption{Submodularity example with sensor placements.}
	\end{center}
\end{figure}

This basically means that submodular functions exhibit a diminishing return property which makes them particularly interesting for optimization. A very intuitive example is the optimum placement of sensors in an area (see figure \ref{fig:submodular}). Sensors can be placed on a grid of locations and function F computes the surface of the space that is being sensed. Note first that adding a new sensor i to the current set S cannot decrease the value of F : $F(S+\{i\})>=F(S)$. Furthermore, if we add sensor i to a small set $S_1$ we tend to get larger improvements than if we add i to a larger set $S_2 \supset S_1$.

Generally speaking, finding the set $ S_k^{\star} $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only a feasible solution on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}. This heuristic starts with a set S (possibly empty) and iteratively adds the element i that exhibits the highest marginal gain : $ F(S \cup \{i\}) \geq F(S \cup \{j\})\ \forall j $.

@@

\begin{figure}
\includegraphics[scale=.55]{plot5.pdf}%./figure_2/figure_2.pdf}
\caption{At $ t_k=800 $, $P_5 $ goes from $-1$ to $-2$ (power inbalance). Consequently, the frequencies $ \dot{ \theta }_i $ deviate from the synchronized state. At $ t_k + d = 1000$, optimal control inputs (see equation \ref{eq:ustar}) are injected at nodes 0, 2, and 4 (nodes with ellipses) such that the system is brought to the synchronized state at time $ t_k + d + T = 1500 $ (T is the control time).}

@@

The purpose of selecting the drivers according to a Gramian related metric is to minimize, on average, the amount of control energy needed. Conversely, if we select the drivers randomly, we expect to need, on average, more energy to control the system. In figure \ref{fig:densities}, we compare the average control energy  $log(\mathcal{E})$ in function of the proportion of drivers amoung the total number of nodes $n_D = N_D/N_{nodes}$, for a set of drivers selected thanks to our algorithm with randomly selected drivers (in both cases overloading and battery limits constraints are satisfied). We draw $10^4$ scale-free topologies with 50 nodes, and random power distributions and line capacities.  For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $. One is chosen randomly and the other is found with algorithm \ref{algo_1}. For both sets, if the control is possible, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$. Since the algorithm is based on $ Tr[W^{-1}]$, we could plot $ Tr[W^{-1}]$ in function of $n_D$ but we prefer to validate our method by computing the actual average energy. As expected, both curves decrease as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease, but we tend to use less energy when the drivers are selected with our algorithm. Note that this difference tends to zero when $n_D$ tends to one, because almost all nodes are then selected, yielding little flexibility for optimization.

@@

We investigate next how the topology of the grid and the physical constraints affect the minimum size of the driver set. The driver set size $n_D$ is no longer selected randomly, but minimized. We consider the simple case of an Erdös-Rényi topology with probability of connection p. We show on Figure \ref{fig:erdos_renyi} how the minimum size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :

>>>>

We study the placement of storage in networks of prosumers that may both consume and produce electricity using renewable generators (DER). In this stochastic context, power imbalances are very likely, such that the system might lose synchrony and require control to bring it back to the synchronized state. We consider the control inputs as being the amount of power injected in the grid at the chosed storage locations. In a context where generators and loads are stochastic and not fixed, the optimal placement of the storages beforehand is not trivial. The dynamic is written as a system of differential equations and the physical constraints of the system are included. We then propose and validate an algorithm based on a submodular optimization that minimizes the average energy required to control the system.

%We first design a model and simplify the underlying dynamic of such network using a second order coupled oscillators network model. Under some conditions, the system synchronizes to a common frequency. However, in case of perturbations in the power distribution, the system might lose synchrony and require control to bring it back to the stable state. In this situation, control can be seen as energy  absorbed or injected in the system at specific locations. Moreover, the power outputs of the prosumers are susceptible to change such that loads and generators are not fixed. In this context, it is important to select the right subset of nodes that yields, on average, the cheapest control in terms of energy. We propose and validate an algorithm based on a submodular optimization for chosing the driver nodes.
\end{abstract}

@@

The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems. Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility. Within these systems, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication. Agents responsible for generation and load portfolio managment, the so-called prosumers \cite{Parag2016} because they may both produce and consume electricity, will provide services (generation, load shedding, frequency regulation) against remuneration. Insure grid stability within this uncertain context appears as a complex task.

%Several studies highlighted the need for increasing the storage capacity of the system. Some of them even consider electric vehicles as moving capacities that could be used for frequency regulation. In the case of fixed storage devices, the locations where they should be installed is an important question since it impacts the performance of the system. In a very static and top-down architecture, where loads and generators are fixed, this question is far from trivial. But in a smart grid scenario with bi-directional flows and agents that produce and consume depending on weather conditions, this becomes even more challenging.

In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes. In such a scenario with a high penetration of renewables, perturbations in the power distributions of the prosumers are very likely. Re-balancing the production and the consumption is then necessary and requires that we have the capability to control the grid's dynamic. In \cite{Gkatzikis2015}, the authors study the placement of storage equipments within microgrids in a collaborative scheme. The authors solve an interesting optimization problem mixing placement-dimensioning of storage with utilization. Moreover, \cite{Gkatzikis2015} uses a Nash bargaining framework to determine how prosumers should share the costs and benefits of the storages. However, \cite{Gkatzikis2015} does not take into account the grid dynamic as well as its controllability. Control of smart grid systems has been recently studied in \cite{Farraj2015}, where the authors use linear-quadratic optimal control theory in order to compute DER outputs. Control inputs rely on information collected by sensors and phasor measurement units and a communication network is used for sharing the information. The performance is then studied under practical limitations such as latency, sampling rate, or signal-to-noise ratio. The objective of \cite{Farraj2015} is the control of the system when perturbations occur. The control signals are applied to the DER, meaning that \cite{Farraj2015} does not consider a system where storage would be present and used for control.

@@

The present work differs from \cite{Gkatzikis2015} and \cite{Farraj2015} in that we study the placement of the storages from a controllability perspective in a stochastic context due to DER. Moreover, we aim at models and methods that could be scalable to some extent.

We are given a prosumer network and we have to find both the number and locations of the storages that should be deployed. Once this choice is made, we have to stick with it whatever perturbations may occur. We are looking for the smallest set of storages, as well as their locations, that will require small control energy. Since we do not have any prior on the future perturbations, we minimize the average control energy required to move the system around the state space.

@@

The power grid dynamic is represented by a coupled oscillator network \cite{Filatrella2008}. Each node, representing a prosumer, is modeled as an oscillator with a phase angle and a natural frequency that depends on its production/consumption. The oscillators are interconnected such that their frequencies are impacted by the ones of their neighbors. As we will see, the phase angles dynamics can be written as a system of coupled second order differential equations \cite{Filatrella2008}.

@@

In this section, we introduce some key notions related to the controllability of complex networks. We consider a graph $G=(V,E)$ of N nodes, where V is the vertex set and E is the edge set. The topology of this network can be encoded through the adjacency matrix M, which element $m_{ij} \in \{0,1\}$ indicates whether i and j are connected. If the edges are weighted we can set $m_{ij} = w_{ij}$ where $w_{ij}$ is the weight of edge$(i,j)$, and zero if there is no edge. Graphs are often used to model complex interactions inside a population. Let us attribute to each node i the variable $ x_i $ that represents the state of node i. Depending on the model, this state can represent an opinion, a physical quantity, a probability and so on. If there is an edge between i and j, we might expect some interaction that would eventually change the values of $x_i$ and $x_j$. Such a dynamic is usually described with a system of differential equations. If we assume a linear first order dynamic, this can be written as $ \dot{X}(t) = A X(t) $, where $X(t)=\{x_0(t),...,x_{N-1}(t)\}$ is the vector of node states at time t, and matrix A encodes both the dynamics and the topology of the system (A can be, among other, the weighted adjacency matrix of the underlying graph).

If there is no action from the outside, the system state will evolve according to $ X(t) = X_0 e^{At} $ where $X_0$ is the initial state of the system. Imagine now, that this dynamic could be somehow influenced by injecting some signals. More precisely, let us assume that we can inject these signals in a subset of the nodes, called drivers, such that the dynamics becomes $ \dot{ X }(t) = A X(t) + B u(t) $, where the matrix B indicates which nodes receive the signals and $ u(t) $ is the vector of inputs at time t. Assume that the system is in some initial state $X_0$ and we aim at bringing it to some final state $X_T$ in some amount of time T. Control can be seen as finding the sequence of inputs $\{u(t_0),...u(t_T)\} $ that would do it given the dynamics A. Furthermore, we say that the system is controllable in T steps if it can be steered from any initial state $ X_0 $ to any final state $ X_T $ through a sequence of control inputs.

It was shown by Kalman that a linear system $(A,B)$ is controllable if the controllability matrix $ C = [B,AB,A^2B,...,A^{N-1}B] $ has full rank, that is, $rank(C) = N$. Nevertheless, when studying the controllability of non trivial networks, using the Kalman's criterion is not a simple task since matrix C becomes huge. Furthermore numeric precision becomes quickly an issue when computing the rank of C. A different approach to study controllability in complex network was introduced later on by Lin \cite{1100557}. This relates the search of the minimum driver set to a maximum matching in a bipartite graph, which can be done with the Hopcroft-Karp algorithm in $ O(|E|\sqrt{|V|}) $ in the worst case. This requires that the system $(A,B)$ is a structured system, meaning that all elements in A and B are either fixed zeros or independant free parameters. In such a situation, structural control is extremely powerfull since it allows to discuss network control easily even if the exact weights of the edges are unknown. Note however that the independance of the parameters is a key assumption. If A is the adjacency matrix of an undirected network for instance, independance is not verified (since A is symmetric). In this case, one should use different tools to study controllability \cite{Yuan2014}.

The fact that a system is controllable or not is not the only criterion of interest. Indeed, if it is controllable, the amount of control energy required also appears as a key criterion. Actually, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $X_0$ to $X_T$. Among all these possibilities, optimal control is devoted to find the sequence that minimizes some cost function that can depend on the states of the system $X(t)$, the control inputs $u(t)$, and the final state $X_T$. In the following, these control inputs will be used to model the actions of storage devices in a power grid. We are therefore particularly interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}. Let $ \mathcal{E} = \int_{0}^{T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system. It can be shown \cite{Liu2015} that the control inputs that minimize $ \mathcal{E} $ can be written as :

@@

\noindent where $ \nu_f = X_T - X_0 e^{AT} $ is the difference between the desired final state $X_T$ and the final free state $X_0 e^{AT} $, and $W(t) = \int_{0}^{t} e^{A \tau}BB^{T} e^{A^{T} \tau} d\tau $ is called the Gramian matrix of the system. It can also be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace. Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :

@@

\noindent where $ K_{Di} $ is the dissipation constant of entity i. Furthermore, it also accumulates kinetic energy at a rate :

@@

\noindent where $ I_{i}$ is the moment of inertia of entity i. For simplicity, we consider that all entities have the same dissipation constants($K_D$) and moment of inertia (I).

@@

\noindent with $ P_{ij}^{MAX} $ being the maximum capacity of the line $(i,j)$. Each entity i is then described by a power balance equation of the type :

@@

\noindent where $ P_{S,i} $ is the power of an ideal source or sink at node i. By substituting equations \ref{eq:diss}, \ref{eq:acc}, and \ref{eq:trans} into equation \ref{eq:bilan} and re-arranging the terms, we obtain the following non-linear coupled system of equations :

@@

\noindent where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $. In order not to overload the equations, we simplify the constant term $ \frac{K_D \Omega}{I} $ by working in a rotating frame such that $ \psi_i=\frac{P_{S,i}}{I \Omega} $.

@@

\noindent With transition matrix A :

@@

On the other hand, the power that flows on line $ (i,j) \in E$ can be written as $ P_{i \longrightarrow j} = -P_{ij}^{MAX} sin\left( \theta_j(t) - \theta_i(t) \right) $. If $ \theta_j(t) - \theta_i(t) $ is small, the flow constraints at time t are straightforward :

@@

Constraints \ref{eq:flow_cons} \ref{synchro_constraint} and \ref{zero_sum_constraint} ensure that the system is able to synchronize at $\Omega$ without overloading lines.

@@

\noindent where $ \Lambda(0) $ is the vector of initial levels in the batteries. Obviously, $\Lambda(t)$ has to stay within possible bounds : $ \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{MAX} $. Which can be written in terms of the control inputs :

@@

Recall that $\mathcal{E}_{min}$ (see eq. \ref{eq:energy}) depends on the initial and final states as well as on the inverse of the gramian matrix W, which only depends on A and B. W can thus be used to obtain information about the average control energy required to move the system in the state space. In the prosumer network that we consider, generators and loads ($\Psi$) are suceptible to change, meaning that initial ($X_0$) and final ($X_f$) states might also vary. In such a scenario, we prefer to aim at good performance on average, rather than very good performance in few specific cases and bad performance in all other situations.

@@

%\begin{figure}
%	\label{fig:submodular}
%	\begin{center}
%	\includegraphics[scale=.25]{./figure_1/figure_1}
%	\caption{Submodularity example with sensor placements.}
%	\end{center}
%\end{figure}

%A very intuitive example is the optimum placement of sensors in an area (see figure \ref{fig:submodular}). Sensors can be placed on a grid of locations and function F computes the surface of the space that is being sensed. Note first that adding a new sensor i to the current set S cannot decrease the value of F : $F(S+\{i\})>=F(S)$. Furthermore, if we add sensor i to a small set $S_1$ we tend to get larger improvements than if we add i to a larger set $S_2 \supset S_1$.

This basically means that submodular functions exhibit a diminishing return property which makes them particularly interesting for optimization. Generally speaking, finding the set $ S_k^{\star} $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only feasible on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}. This heuristic starts with a set S (possibly empty) and iteratively adds the element i that exhibits the highest marginal gain : $ F(S \cup \{i\}) \geq F(S \cup \{j\})\ \forall j $.

@@

\begin{figure}
\includegraphics[scale=.55]{plot5.pdf}%./figure_2/figure_2.pdf}
\caption{At $ t=800 $, $P_5 $ goes from $-1$ to $-2$ (power inbalance). Consequently, the frequencies $ \dot{ \theta }_i $ deviate from the synchronized state. At $ t = 1000$, optimal control inputs (see equation \ref{eq:ustar}) are injected at nodes 0, 2, and 4 (nodes with ellipses) such that the system is brought to the synchronized state at time $ t = 1500 $ (control time $ T = 500 $).}

@@

The purpose of selecting the drivers according to a Gramian related metric is to minimize, on average, the amount of control energy needed. Conversely, if we select the drivers randomly, we expect to need, on average, more energy to control the system. In figure \ref{fig:densities}, we compare the average control energy  $\mathcal{E}$ in function of the proportion of drivers among the total number of nodes $n_D = N_D/N_{nodes}$, for a set of drivers selected thanks to our algorithm with randomly selected drivers (in both cases overloading and battery limits constraints are satisfied). We draw $10^4$ scale-free topologies with 50 nodes, and random power distributions and line capacities.  For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $. One is chosen randomly and the other is found with algorithm \ref{algo_1}. For both sets, if the control is possible, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$. Since the algorithm is based on $ Tr[W^{-1}]$, we could plot $ Tr[W^{-1}]$ in function of $n_D$ but we prefer to validate our method by computing the actual average energy. As expected, both curves decrease as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease, but we tend to use less energy when the drivers are selected with our algorithm. Note that this difference tends to zero when $n_D$ tends to one, because almost all nodes are then selected, yielding little flexibility for optimization.

@@

We investigate next how the topology of the grid and the physical constraints affect the minimum size of the driver set. $n_D$ is no longer selected randomly, but minimized. We consider the simple case of an Erdös-Rényi topology with probability of connection p. We show on Figure \ref{fig:erdos_renyi} how the minimum size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :
